Tail associations in ecological variables and their impact on extinction risk

### 2020-03-27 MainText.Rmd
minor edit

<!--stochastic matrix model : age-structured-->
To investigate Q1 from the Introduction,
we considered a simple stochastic age-structured model
with two age classes, 
\begin{equation}
\left({\begin{array}{cc} E(t+1) \\ A(t+1) \end{array}}\right) = 
\left(\begin{array}{cc} 0 & f_a(t)\\ s_e(t) & 0 \end{array}\right)
\left(\begin{array}{c} E(t) \\ A(t) \end{array}\right), \label{eq_age_str}
\end{equation}
\noindent where $E(t)$ and $A(t)$ are the numbers or population densities of eggs 
and adults, respectively, at time $t$. We chose a very simple model
because our goal was to carry out an initial evaluation, only, of 
the realism of a potential mechanism. The stochastic vital 
rates $s_e(t)$ and $f_a(t)$ in Eq. \ref{eq_age_str} were selected, independently 
across times, as follows. The egg survival rate $s_e(t)$ was 
drawn from a beta distribution with parameters $\alpha=3$ and $\beta=2$. 
Adult fecundity was $f_a(t)=\tilde{f}_a(t)\exp(-A(t)/\Omega)$, where $\tilde{f}_a(t)$
was drawn from a gamma distribution with shape and scale parameters $k=2$
and $\theta=1$, respectively, 
and $\Omega$ is a density dependence parameter. Allowing $\Omega$ to tend to $\infty$
$\exp(-A(t)/\Omega)=1$ so that the model did not include density dependence
(it was then a standard stochastic matrix model). Finite, positive values of 
$\Omega$ made the model nonlinear, with density dependence in the adult
fecundity rate. Stochastic matrix models are very commonly used in
evolutionary ecology and quantitative conservation biology, so our model
is an example of a very widely used category of models. See, for instance,
the classic books of @Caswell2000, and @MorrisDoak2002.

>>>>

<!--stochastic matrix model : age-structured-->
To investigate Q1 from the Introduction,
we considered a simple stochastic age-structured model
with two age classes, 
\begin{equation}
\left({\begin{array}{cc} E(t+1) \\ A(t+1) \end{array}}\right) = 
\left(\begin{array}{cc} 0 & f_a(t)\\ s_e(t) & 0 \end{array}\right)
\left(\begin{array}{c} E(t) \\ A(t) \end{array}\right), \label{eq_age_str}
\end{equation}
\noindent where $E(t)$ and $A(t)$ are the numbers or population densities of eggs 
and adults, respectively, at time $t$. We chose a very simple model
because our goal was to carry out an initial evaluation, only, of 
the realism of a potential mechanism. The stochastic vital 
rates $s_e(t)$ and $f_a(t)$ in Eq. \ref{eq_age_str} were selected, independently 
across times, as follows. The egg survival rate $s_e(t)$ was 
drawn from a beta distribution with parameters $\alpha=3$ and $\beta=2$. 
Adult fecundity was $f_a(t)=\tilde{f}_a(t)\exp(-A(t)/\Omega)$, where $\tilde{f}_a(t)$
was drawn from a gamma distribution with shape and scale parameters $k=2$
and $\theta=1$, respectively, 
and $\Omega$ is a density dependence parameter. Allowing $\Omega$ to tend to $\infty$,
$\exp(-A(t)/\Omega) \rightarrow 1$, so that, in that limit, the model did not include density dependence
(it was then a standard stochastic matrix model). Finite, positive values of 
$\Omega$ made the model nonlinear, with density dependence in the adult
fecundity rate. Stochastic matrix models are very commonly used in
evolutionary ecology and quantitative conservation biology, so our model
is an example of a very widely used category of models. See, for instance,
the classic books of @Caswell2000, and @MorrisDoak2002.
